\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-27 paper.qmd
Added model initial projections (need to work on this)

@@

@@

distance_df_small_edges_two_curvy2 <- distance_two_curvy |>
  filter(distance < benchmark_two_curvy)

tr_from_to_df_two_curvy4 <- inner_join(
  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy2,
  by = c("from", "to"))

@@

@@

@@

<!-- langevitour with pD model for S-curve -->
::: {#fig-two_curvy-sc-best layout-ncol="4" fig-pos="H"}
![](figures/scurve/tsne_trimesh_layout.png)

![](figures/scurve/sc_tsne_best_1.png)

![](figures/scurve/sc_tsne_best_2.png)

![](figures/scurve/sc_tsne_best_3.png)

Model in \gD{}, on the tSNE layout, and three views of the fit in projections from $7\text{-}D$, for the S-curve data ($(s_1, \ s_2) = (-0.170, \ -0.137)$, $b = 256 \  (16, \ 16)$, $m = 70$, benchmark value to remove low density hexagons is $0.250$, and benchmark value to remove large edges is $0.180$). MSE is $0.0432$. The model closely fits the shape, but it has some twists (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/ZuLdp89qJ6g>).
:::

>>>>

df_b_two_curvy2 <- df_bin_two_curvy2 |>
  dplyr::filter(hb_id %in% df_bin_centroids2$hexID) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b_two_curvy2 <- df_b_two_curvy2[match(df_bin_centroids2$hexID, df_b_two_curvy2$hb_id),] |>
  dplyr::select(-hb_id)

@@

trimesh_two_curvy_removed1_with_data <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy3,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#33a02c") +
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#000000",
            alpha = 0.5) +
  interior_annotation("b")

@@

distance_df_small_edges_two_curvy3 <- distance_two_curvy |>
  filter(distance < benchmark_two_curvy)

tr_from_to_df_two_curvy4 <- inner_join(
  tr_from_to_df_two_curvy2, distance_df_small_edges_two_curvy3,
  by = c("from", "to"))

@@

```{r}
#| label: best-fit-two-curvy-proj

# Apply the scaling
df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_b_two_curvy2)

scaled_two_curvy_data_true_model <- scaled_two_curvy |>
  filter(type == "true model") |>
  select(-type)

scaled_two_curvy_data_model <- scaled_two_curvy |>
  filter(type == "model") |>
  select(-type)

projected_df <- projected |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  dplyr::mutate(ID = dplyr::row_number())

## For fitted model
projected_model <- as.matrix(scaled_two_curvy_data_model) %*% projection_scaled

model_df <- dplyr::left_join(
  distance_df_small_edges_two_curvy2 |> select(-distance), 
  projected_model_df, 
  by = c("from" = "ID"))

## For true model

projected_true_model <- as.matrix(scaled_two_curvy_data_true_model) %*% projection_scaled

projected_true_model_df <- projected_true_model |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  dplyr::mutate(ID = dplyr::row_number())

true_model_df_proj <- dplyr::left_join(
  wireframe_true_model, 
  projected_true_model_df, 
  by = c("from" = "ID"))

names(true_model_df_proj)[3:NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_from")

true_model_df_proj <- dplyr::left_join(true_model_df_proj, projected_true_model_df, by = c("to" = "ID"))
names(true_model_df_proj)[(2 + NCOL(projected_true_model_df)):NCOL(true_model_df_proj)] <- paste0(names(projected_true_model_df)[-NCOL(projected_true_model_df)], "_to")

axes_obj <- gen_axes(  
  proj = projection * 5,
  limits = 0.15,
  axis_pos_x = -0.1,
  axis_pos_y = -0.1,
  axis_labels = names(scaled_two_curvy_data),
  threshold = 0)

two_curvy_proj_umap_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#253494") +
  geom_point(
    #size = 0.2,
    alpha = 0.3,
    color = "#969696") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("c") +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_all_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = true_model_df_proj, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#253494") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#253494") +
  geom_point(
    #size = 0.2,
    alpha = 0.3,
    color = "#969696") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("d") +
  theme(
    legend.position = "none"
  )

@@

<!-- langevitour with pD model for two-curvy clusters -->
```{r}
#| label: best-fit
#| fig-cap: "Best fit"
#| fig-width: 10
#| fig-height: 10

nldr_two_curvy + 
  trimesh_two_curvy_removed1_with_data +
  two_curvy_proj_umap_model +
  two_curvy_proj_umap_all_model +
  plot_layout(ncol = 2)
```

@@

### 2024-09-27 paper.tex
Added error plot
Added long edge plots
Added model initial projections (need to work on this)

@@

@@

@@

\begin{minipage}{0.25\linewidth}
\includegraphics{figures/scurve/tsne_trimesh_layout.png}\end{minipage}%
%
\begin{minipage}{0.25\linewidth}
\includegraphics{figures/scurve/sc_tsne_best_1.png}\end{minipage}%
%
\begin{minipage}{0.25\linewidth}
\includegraphics{figures/scurve/sc_tsne_best_2.png}\end{minipage}%
%
\begin{minipage}{0.25\linewidth}
\includegraphics{figures/scurve/sc_tsne_best_3.png}\end{minipage}%

\caption{\label{fig-two_curvy-sc-best}Model in \gD{}, on the tSNE
layout, and three views of the fit in projections from \(7\text{-}D\),
for the S-curve data (\((s_1, \ s_2) = (-0.170, \ -0.137)\),
\(b = 256 \  (16, \ 16)\), \(m = 70\), benchmark value to remove low
density hexagons is \(0.250\), and benchmark value to remove large edges
is \(0.180\)). MSE is \(0.0432\). The model closely fits the shape, but
it has some twists (The \textbf{langevitour} software is used to view
the data with a tour, and the full video is available at
\url{https://youtu.be/ZuLdp89qJ6g}).}

>>>>

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/best-fit-1.pdf}

\caption{Best fit}

@@

\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}

\caption{\label{fig-p-d-error-in-2d-two-curvy}error}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-34-1.pdf}

\caption{long edge removal}

@@
